---
title: "IST687_Group2_Final"
author: "William Jeffery, Damon Porter, Kelly Schillaci, Jacob Marsh"
date: "December 16, 2018"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

The following code chunks are the different portions of Group 3's analysis on the gaming dataset. The team originally set out to answer three questions that were slightly modified as time went on. The team's ending questions are:

What is the popularity by genre?

What are the top selling games by company and region?

What are the top selling games by platform and genre?

The answers to those three questions and required analysis will be answered with the below code chunks.

This first chunk of code attempts to apply linear models to the data to see if we are able to perdict marketing trends using the different sales values in the data.

```{r final}

# Import Data from ratings.csv

a <- read.csv("C:/Users/kdoyl/OneDrive/Documents/IST687/ratings.csv", 
              header=TRUE, sep=",", na.strings="NA", dec=".", strip.white=TRUE)

# Import Data from main dataset

b <- read.csv("C:/Users/kdoyl/OneDrive/Documents/IST687/newVG.csv", 
              header=TRUE, sep=",", na.strings="NA", dec=".", strip.white=TRUE)

# Assign names to a

names(a) <- c("Name","Year","Release Date","Platform","Score") 

# Merge Datasets
total <- merge(a,b, by =c("Name","Platform"), na.rm=TRUE)

# Create function to numberize values
Number <- function(inputVector)
{return(as.numeric(inputVector))}

# run Number function on dataset
total$Global_Sales <- Number(total$Global_Sales)
total$NA_Sales <- Number(total$NA_Sales)
total$EU_Sales <- Number(total$EU_Sales)
total$JP_Sales <- Number(total$JP_Sales)
total$Other_Sales <- Number(total$Other_Sales)

# Bring in ggplot2 and data.table
library(ggplot2)
library(data.table)


#Sum of Total Sales per year
DT <- data.table(total)
DT <- DT[, sum(Global_Sales), by = Year.y]
plot(DT)

# The following code tests the dataset using several different linear
# models
model1 <- lm(formula=Global_Sales ~ Score + Genre, data=total)
summary(model1)

model2 <- lm(formula=Global_Sales ~ Score + Platform, data=total)
summary(model2)

# Of note
model1NA <- lm(formula=NA_Sales ~ Score + Genre, data=total)
summary(model1NA)

model2NA <- lm(formula=NA_Sales ~ Score + Platform, data=total)
summary(model2NA)

# Of note
model1EU <- lm(formula=EU_Sales ~ Score + Genre, data=total)
summary(model1EU)

model2EU <- lm(formula=EU_Sales ~ Score + Platform, data=total)
summary(model2EU)

model1JP <- lm(formula=JP_Sales ~ Score + Genre, data=total)
summary(model1JP)

# Of note
model2JP <- lm(formula=JP_Sales ~ Score + Platform, data=total)
summary(model2JP)

# Of note
model1Other <- lm(formula=Other_Sales ~ Score + Genre, data=total)
summary(model1Other)

model2OTher <- lm(formula=Other_Sales ~ Score + Platform, data=total)
summary(model2OTher)


```

This chunk of code attempts to display the answer to the three questions listed above.

The answer to this question makes use of the sqldf function to extract several different smaller data frames to be passed to the plotting function later on. The bar charts shows the top sales in millions every 10 years starting at the beginning of the team's data.

Finally, the heat maps at the end show the top selling Genres by years for the available data.

```{r things}

library(sqldf)
library(lubridate)
#install.packages("lattice")
library(lattice)
library(plyr)
library(reshape2)
library(dplyr)
library(ggplot2)
install.packages("wesanderson")
library(wesanderson)

vgSales <- read.csv("C:/Users/kdoyl/OneDrive/Documents/IST687/newVG.csv")
vgRatings <- read.csv("C:/Users/kdoyl/OneDrive/Documents/IST687/ratings.csv")
vgSales$Year <- as.Date(vgSales$Year,"%Y")

vgSales$Year <- year(as.Date(vgSales$Year, format = "%Y-%m-%d"))
vgSalesRating <- merge(vgSales,vgRatings,by.x = c("Name","Platform","Year"), by.y =c("NAME","PLATFORM", "Release_Year") )

#################################################################################################################################

#Extract data using sqldf and SQL Syntax
genreAll <- sqldf("SELECT Genre, SUM(Global_Sales) AS TotalSales
                  FROM vgSales 
                  GROUP BY GENRE
                  ORDER BY GENRE ASC")

genreTopGamesAll <- sqldf("SELECT *
                          FROM vgSales 
                          ORDER BY Global_Sales DESC
                          LIMIT 100")

genre80s <- sqldf("SELECT Genre, SUM(Global_Sales) AS TotalSales
                  FROM vgSales 
                  WHERE Year BETWEEN 1980 and 1989
                  GROUP BY GENRE
                  ORDER BY GENRE ASC")

genreTopGames80s <- sqldf("SELECT *
                          FROM vgSales 
                          WHERE Year BETWEEN 1980 and 1989
                          ORDER BY Global_Sales DESC
                          LIMIT 100")

genre90s <- sqldf("SELECT Genre, SUM(Global_Sales) AS TotalSales
                  FROM vgSales 
                  WHERE Year BETWEEN 1990 and 1999
                  GROUP BY GENRE
                  ORDER BY GENRE ASC")

genreTopGames90s <- sqldf("SELECT *
                          FROM vgSales 
                          WHERE Year BETWEEN 1990 and 1999
                          LIMIT 100")

genre2000s <- sqldf("SELECT Genre, SUM(Global_Sales) AS TotalSales
                  FROM vgSales 
                  WHERE Year BETWEEN 2000 and 2009
                  GROUP BY GENRE
                  ORDER BY GENRE ASC")

genreTopGames2000s <- sqldf("SELECT *
                  FROM vgSales 
                  WHERE Year BETWEEN 2000 and 2009
                  LIMIT 100")

genre2010s <- sqldf("SELECT Genre, SUM(Global_Sales) AS TotalSales
                  FROM vgSales 
                  WHERE Year BETWEEN 2010 and 2019
                  GROUP BY GENRE
                  ORDER BY GENRE ASC")

genreTopGames2010s <- sqldf("SELECT *
                  FROM vgSales 
                  WHERE Year BETWEEN 2010 and 2019
                  LIMIT 100")

eightiesGenreRegion <- sqldf("SELECT Genre, SUM(NA_Sales) as Total_NA_Sales, SUM(JP_Sales) AS Total_Japanese_Sales,SUM(Other_Sales) AS Total_Other_Sales,SUM(EU_Sales) AS Total_European_Sales
                             FROM vgSales
                             WHERE YEAR BETWEEN 1980 AND 1989
                             GROUP BY Genre")

ninteiesGenreRegion <- sqldf("SELECT Genre, SUM(NA_Sales) as Total_NA_Sales, SUM(JP_Sales) AS Total_Japanese_Sales,SUM(Other_Sales) AS Total_Other_Sales,SUM(EU_Sales) AS Total_European_Sales
                             FROM vgSales
                             WHERE YEAR BETWEEN 1990 AND 1999
                             GROUP BY Genre")

GenreRegion2000s <- sqldf("SELECT Genre, SUM(NA_Sales) as Total_NA_Sales, SUM(JP_Sales) AS Total_Japanese_Sales,SUM(Other_Sales) AS Total_Other_Sales,SUM(EU_Sales) AS Total_European_Sales
                             FROM vgSales
                             WHERE YEAR BETWEEN 2000 AND 2009
                             GROUP BY Genre")

GenreRegion2010s <- sqldf("SELECT Genre, SUM(NA_Sales) as Total_NA_Sales, SUM(JP_Sales) AS Total_Japanese_Sales,SUM(Other_Sales) AS Total_Other_Sales,SUM(EU_Sales) AS Total_European_Sales
                             FROM vgSales
                             WHERE YEAR BETWEEN 2010 AND 2019
                             GROUP BY Genre")

#################################################################################################################################

# Create Bar charts to display Top sellings games by genre
BarTotal80s <- ggplot(data = genre80s, aes(x= Genre, y = TotalSales )) + geom_col(aes(size = TotalSales, fill = Genre))
BarTop1980s <- ggplot(data = genreTopGames80s, aes(x=Name, y= Global_Sales)) + geom_col(aes(fill = Genre ),na.rm = TRUE)
BarTop1980s <- BarTop1980s + theme(axis.text.x  = element_text(angle = -90, hjust = 0))
BarTop1980s <- BarTop1980s + ylab("Sales in Millions")
BarTop1980s <- BarTop1980s + ggtitle("The 100 Top Selling Games of The 1980s by Genre")

BarTotal90s <- ggplot(data = genre90s, aes(x= Genre, y = TotalSales )) + geom_col(aes(size = TotalSales, fill = Genre))
BarTop1990s <- ggplot(data = genreTopGames90s, aes(x=Name, y= Global_Sales)) + geom_col(aes(fill = Genre ),na.rm = TRUE)
BarTop1990s <- BarTop1990s + theme(axis.text.x  = element_text(angle = -90, hjust = 0))
BarTop1990s <- BarTop1990s + ylab("Sales in Millions")
BarTop1990s <- BarTop1990s + ggtitle("The 100 Top Selling Games of The 1990s by Genre")

BarTotal2000s <- ggplot(data = genre2000s, aes(x= Genre, y = TotalSales )) + geom_col(aes(size = TotalSales, fill = Genre))
BarTop2000s <- ggplot(data = genreTopGames2000s, aes(x=Name, y= Global_Sales)) + geom_col(aes(fill = Genre ),na.rm = TRUE)
BarTop2000s <- BarTop2000s + theme(axis.text.x  = element_text(angle = -90, hjust = 0))
BarTop2000s <- BarTop2000s + ylab("Sales in Millions")
BarTop2000s <- BarTop2000s + ggtitle("The 100 Top Selling Games of The 2000s by Genre")

BarTotal2010s <- ggplot(data = genre2010s, aes(x= Genre, y = TotalSales )) + geom_col(aes(size = TotalSales, fill = Genre))
BarTop2010s <- ggplot(data = genreTopGames2010s, aes(x=Name, y= Global_Sales)) + geom_col(aes(fill = Genre ),na.rm = TRUE)
BarTop2010s <- BarTop2010s + theme(axis.text.x  = element_text(angle = -90, hjust = 0))
BarTop2010s <- BarTop2010s + ylab("Sales in Millions")
BarTop2010s <- BarTop2010s + ggtitle("The 100 Top Selling Games of The 2010s by Genre")

BarTotalAll <- ggplot(data = genreAll, aes(x= Genre, y = TotalSales )) + geom_col(aes(size = TotalSales, fill = Genre))
BarAll <- ggplot(data = genreTopGamesAll, aes(x=Name, y= Global_Sales)) + geom_col(aes(fill = Genre ),na.rm = TRUE)
BarAll <- BarAll + theme(axis.text.x  = element_text(angle = -90, hjust = 0))
BarAll <- BarAll + ylab("Sales in Millions")
BarAll <- BarAll + ggtitle("The 100 Top Selling Games of All Time by Genre")

#################################################################################################################################

#Extract Total sales by Genre and convert to from vector to data frame.
GenreNASales <- as.data.frame(tapply(vgSales$NA_Sales, vgSales$Genre, sum))
GenreEUSales <- as.data.frame(tapply(vgSales$EU_Sales, vgSales$Genre, sum))
GenreJPSales <- as.data.frame(tapply(vgSales$JP_Sales, vgSales$Genre, sum))
GenreOtherSales <- as.data.frame(tapply(vgSales$Other_Sales, vgSales$Genre, sum))

#Add new columns for using rownames
GenreNASales$Genre <- rownames(GenreNASales)
GenreEUSales$Genre <- rownames(GenreEUSales)
GenreJPSales$Genre <- rownames(GenreJPSales)
GenreOtherSales$Genre <- rownames(GenreOtherSales)

#rename column names
names(GenreNASales) <- c("North American Sales", "Genre")
names(GenreEUSales) <- c("European Sales", "Genre")
names(GenreJPSales) <- c("Japanese Sales", "Genre")
names(GenreOtherSales) <- c("Other Sales", "Genre")
#################################################################################################################################

# Joining Total Genre Sales for Each Region
allRegions <- join_all(list(GenreNASales,GenreJPSales,GenreEUSales,GenreOtherSales), by='Genre', type = 'full')
allRegions <- allRegions[c(2,1,3,4,5)]
allRegions.m <- melt(allRegions)

# Heat Map of Total Sales by Genre
genreHeatMap <- ggplot(allRegions.m, aes(variable, Genre)) + geom_tile(aes(fill = value), colour = "white") + scale_fill_gradient(low = "lightblue", high = "purple")
genreHeatMap <- genreHeatMap + geom_text(aes(label = value)) + xlab("Region") 
genreHeatMap <- genreHeatMap + guides(fill = guide_legend(title = "Sales in Millions"))


#Heat Map of Sales/Genre in the 1980s/NES Colors
eightiesGenreRegion.m <- melt(eightiesGenreRegion)
eightiesgenreHeatMap <- ggplot(eightiesGenreRegion.m, aes(variable, Genre)) + geom_tile(aes(fill = value), colour = "white") + scale_fill_gradient(low = "grey",high = "red")
eightiesgenreHeatMap <- eightiesgenreHeatMap + geom_text(aes(label = round(value,2)),color = "black") + xlab("Region") 
eightiesgenreHeatMap <- eightiesgenreHeatMap + guides(fill = guide_legend(title = "Sales in Millions"))

#Heat Map of Sales/Genre in the 1990s/SNES Colors
ninteiesGenreRegion.m <- melt(ninteiesGenreRegion)
ninteiesgenreHeatMap <- ggplot(ninteiesGenreRegion.m, aes(variable, Genre)) + geom_tile(aes(fill = value), colour = "white") + scale_fill_gradient(low = "grey", high = "purple")
ninteiesgenreHeatMap <- ninteiesgenreHeatMap + geom_text(aes(label = round(value,2))) + xlab("Region") 
ninteiesgenreHeatMap <- ninteiesgenreHeatMap + guides(fill = guide_legend(title = "Sales in Millions"))

#Heat Map of Sales/Genre in the 2000s/Playstation 2 Colors
GenreRegion2000s.m <- melt(GenreRegion2000s)
GenreRegion2000sHeatMap <- ggplot(GenreRegion2000s.m, aes(variable, Genre)) + geom_tile(aes(fill = value), colour = "white") + scale_fill_gradient(low = "black", high = "blue")
GenreRegion2000sHeatMap <- GenreRegion2000sHeatMap + geom_text(aes(label = round(value,2)),color = "grey") + xlab("Region") 
GenreRegion2000sHeatMap <- GenreRegion2000sHeatMap + guides(fill = guide_legend(title = "Sales in Millions"))

#Heat Map of Sales/Genre in the 2010s
GenreRegion2010s.m <- melt(GenreRegion2010s)
GenreRegion2010sHeatMap <- ggplot(GenreRegion2010s.m, aes(variable, Genre)) + geom_tile(aes(fill = value), colour = "white") + scale_fill_gradient(low = "deepskyblue3", high = "blue")
GenreRegion2010sHeatMap <- GenreRegion2010sHeatMap + geom_text(aes(label = round(value,2)),color = "white") + xlab("Region") 
GenreRegion2010sHeatMap <- GenreRegion2010sHeatMap + guides(fill = guide_legend(title = "Sales in Millions"))

```

Below we see that the best time to release a video game is December of any year for North America. This is likely because shopping increases towards the end of November and continues through December because of the holidays. We also see that the number of sales drops in January and steadily increases as the year goes on.

```{r thirdchunk}

names(total)[4]<-"Release_Date" ##Changed the column name to remove the space

total$Release_Month <- format(as.Date(total$Release_Date), "%y") ##Created new column that has only release month

total$Release_Month <- as.numeric(total$Release_Month) ##Changed column to be a number instead of a factor

modelBestTime <- lm(formula=NA_Sales ~ Release_Month, data=total) ##Create basic comparison model with NA_Sales as Dependent and Release Month as Independent

summary(modelBestTime) ##Optional

test = data.frame(Release_Month=12) ## Setup with December as the option, can change for different months

predict(modelBestTime,test,type="response") ## Dec = 45.04, Nov = 43.24, Oct = 41.44, Sep = 39.64, etc.


```

In conclusion, the graphs above show the highest selling and popularity of genres of video games as well as the highest selling platform.

Finally, we can see that the best time to sell any game in North America is in November and December of any year. This will allow for the highest sales.